---
title: "The Heterogeneous Impact of Fiscal Policy on Russian Regions"
author: "Ilya Gulenkov"
institute: |
  | International Laboratory for Macroeconomic Analysis
  | Department of Theoretical Economics
  | FES HSE
date: "April 2025"
output: beamer_presentation
fontsize: 10pt
lang: en
header-includes:
  - \usepackage{tikz} 
  - \usepackage[T2A]{fontenc}
  - \usepackage{booktabs}
bibliography: ../Literature/RFS.bib
biblio-style: apalike
---

```{r Setup, echo = F, include = FALSE}
opts <- list(template = "../Technical/tikz2pdf-cyr.tex")
engine <- 'tikz'
knitr::opts_chunk$set(echo = F, message = F, warning = F)
trace(grDevices:::png, quote({
  if (missing(type) && missing(antialias)) {
    type <- "cairo-png"
    antialias <- "subpixel"
  }
}), print = FALSE)
```

```{r Data, echo = F}
set.seed(123)
library(tidyverse)
library(readxl)
library(ggtext)
library(geofacet)
library(scales)
library(ggpattern)
library(ggridges)
library(ggbreak) 
library(spatstat)
library(sandwich)
library(lmtest)
library(kableExtra)
library(ggcorrplot)
library(cluster)    # clustering algorithms
library(flextable)
library(huxtable)
library(stargazer)
options(scipen=9999)
set_flextable_defaults(font.family = 'Times New Roman')
tabForm <- function(x, cap) {
  x %>%
  autofit() %>%
  set_table_properties(layout = "autofit") %>%
  flextable::font(fontname = 'Times New Roman', part = 'all') %>%
  padding(padding = 0, part = "all") %>%
  flextable::align(align = "center", part = "all") %>%
  flextable::align(align = 'justify', part = 'footer') %>%
  colformat_double(big.mark = ",", digits = 2, na_str = "N/A")
  # %>%
  # flextable::set_caption(cap, align_with_table = F, word_stylename = "Table Caption", fp_p = officer::fp_par(text.align = "justify"))
}
figNote <- function(text) {
  officer::fpar(paste0('Примечание: ', text), fp_p=officer::fp_par(text.align = 'justify', padding.top = 0), fp_t = officer::fp_text_lite(font.size = 11))
}
# LOAD STYLING ----
source("../Technical/hse_theme.R")
select <- dplyr::select
hse_colours <- c("#102D69", "#1656A6", "#BFBFBF", "#E61F3D", "#11A0D7", "#FFD746", "#029C63", "#EB681F")
# LOAD DATA ----
QUANTILES <- c(0.15, 0.5, 0.85)
SIG_OPACITY <- 0.75
GRADIENT_QUANTILE_LIMITS <- c(0.15,0.85)
source('../R/3. mult.R', encoding = 'UTF-8')
```

## Motivation

- High regional differentiation in Russia

- Monetary policy transmission depends on regional characteristics [@napalkovRazlichiyaEffektahEdinoy2021;@zverevaVliyanieVnutriregionalnogoNeravenstva2024; @zverevaReakciyaBankovskihStavok2024]:

    - GRP size and structure
    - Local market size
    - Trade geography and intensity
    - Income inequality
    - Employment structure
    - ...
    
- Fiscal policy in a monetary union: countercyclican and redistributive role
    
- The reaction of Russian regions to fiscal policy shocks has not been extensively studied 

## Aim & Contribution

- Aim: estimate the impact of regional government expenditures on Russian regions' economy

- Methodology: global vector autoregression (GVAR), sample period - 2009 - 2024, 79 regions

- Contribution:

    - Fiscal policy effects in Russia: we estimate region-specific multipliers instead of a single national one
    - Heterogeneous effects of macroeconomic policy in a monetary union: multipliers depend on economy size, income inequality and employment structure
    - Cross-regional spillover effects: the size of the effects depends on the size of the source economy and its trade patterns

## Literature: Fiscal Multiplier

- Fiscal multiplier in the U.S. lies between 0.8 and 1.5 [@rameyCanGovernmentPurchases2011]

- The size of the multiplier depends on the business cycle phase [@auerbachFiscalMultipliersRecession2011], the source of economic shocks [@ghassibeStateDependenceFiscal2022], monetary policy reaction and income distribution patterns [@broerFiscalMultipliersHeterogeneousagent2023]

- For Russia estimates vary between 0.1 - 0.9 and depend on the methodology, sample period, budget level and expenditure category [@kudrinByudzhetnayaPolitikaKak2017;@vlasovFiskalnyeMultiplikatoryRossii2018;@zyablickiyOcenkaFiskalnyhMultiplikatorov2020]

## Literature: Local Multiplier and Spatial Effects

- Subnational data allows to estimate local fiscal multipliers [@nakamuraFiscalStimulusMonetary2014;@chodorow-reichGeographicCrossSectionalFiscal2019]:

    - Controls for common country-level factors
    - Higher estimation efficiency due to cross-sectional variation
    - Spatial effects can be studiedd

- Omission of spatial effects can bias regional multiplier estimates downwards [@auerbachLocalFiscalMultipliers2020]

- Spatial links between regions and region clusters in Russia have been extensively documented in the literature [@demidovaGosudarstvennyeRashodyNa2021;@demidovaModeliEkonomicheskogoRosta2016]  

## Model Structure

```{r ModelStructure, fig.cap = "GVAR Model Structure", fig.ext='png', fig.width = 6.5, engine.opts = opts, engine=engine}
\usetikzlibrary{calc,positioning,patterns,angles,intersections,ext.paths.ortho,shapes.multipart,fit,matrix,shadows,shadows.blur,arrows,shapes,arrows.meta,positioning,decorations.pathreplacing,quotes}
\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
            \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}
\newsavebox{\test}
\savebox{\test}{\begin{tikzpicture}[%
    ,node distance=5mm
    ,>=latex'
    ,blockbig/.style = {%
    ,draw
    ,minimum height=10mm,minimum width=28mm
    ,align=center
    }
    ,blocksmall/.style = {%
    ,draw
    ,minimum height=7mm,minimum width=10mm
    ,align=center
    }
    ,blocksmalltext/.style = {%
    ,draw=black!60
    ,minimum height=10mm,minimum width=20mm
    ,align=center,
    font={\footnotesize}
    }
    ,textsmall/.style = {%
    ,minimum height=7mm,minimum width=10mm
    ,align=center,
    font={\footnotesize}
    }
    ,every path/.style={->} 
    ] 
\node[blocksmalltext, anchor = south west] (exp1) {Budget\\spending};
\node[below = 0.1 cm of exp1] (num) {\circled{1}};
%\node[blocksmalltext, right = of exp1] (wage1) {Wages};
\node[blocksmalltext, below left = 0.5 cm and -0.5 cm of exp1] (y1) {Output};
\node[blocksmalltext, below right = 0.5 cm and -0.5 cm of exp1] (infl1) {Inflation};
%\draw[<->] (exp1.east) -- (wage1.west);
%\draw[<->] (wage1.south west) -- (y1.north east);
%\draw[<->] (infl1.east) -- ($(infl1.east) +(0.2,0)$) |- (wage1.east); 
\draw[<->] (y1.east) -- (infl1.west);
\draw[<->] (y1.west) -- ($(y1.west) +(-0.2,0)$) |- (exp1.west);
\draw[<->] (infl1.east) -- ($(infl1.east) +(0.2,0)$) |- (exp1.east); 
\end{tikzpicture}}
\begin{tikzpicture}[%
    ,node distance=5mm
    ,>=latex'
    ,blockbig/.style = {%
    ,draw
    ,minimum height=10mm,minimum width=28mm
    ,align=center
    }
    ,blocksmall/.style = {%
    ,draw
    ,minimum height=7mm,minimum width=10mm
    ,align=center
    }
    ,blocksmalltext/.style = {%
    ,draw=black!60
    ,minimum height=10mm,minimum width=20mm
    ,align=center,
    font={\footnotesize}
    }
    ,textsmall/.style = {%
    ,minimum height=7mm,minimum width=10mm
    ,align=center,
    font={\footnotesize}
    }
    ]
    \node (r1) {\usebox{\test}};
    \node[right = of r1.south east] (r2) {$\dots$};
    \node[right = 1.5 cm of r1] (rN) {\usebox{\test}};
    \draw[decorate,decoration={brace,mirror}] (r1.south west) --
        node[below,align=center] (regname1) {Region 1\\VAR} (r1.south east);
    \draw[decorate,decoration={brace,mirror}] (rN.south west) --
        node[below,align=center] (regnameN) {Region $N$\\VAR} (rN.south east);
    \node[blocksmalltext, below = 0.5 cm of r2] (tradeflows) {Trade flows matrix};
    \node[blocksmalltext, align = center, below = 0.5 cm of tradeflows] (external) {External variables};
    \draw[->] (regname1.south) |- (tradeflows);
    \draw[<-] (tradeflows) -| (regnameN.south);
    \draw[->] (tradeflows.south) -- (external);
    \draw[->] (external.west) -- ($(regname1.west|-external.west) + (-0.25,0)$) node[fill=white] {\circled{2}} |- (regname1.west);
    \draw[->] (external.east) -- ($(regnameN.east|-external.east) + (0.25,0)$) node[fill=white] {\circled{2}} |- (regnameN.east);
    \node[below = 0 cm of external.south] {Spillover effects};
    \draw[thick] ($(r1.north west) +(-0.2,0.2)$) rectangle ($(external.south-|rN.south east) + (0.2,-0.5)$);
    \node[blocksmalltext, above right = 1.1 cm of rN] (urals) {Urals Price};
    \node[blocksmalltext, right = of urals] (fedbudg) {Federal\\budget};
    \node[blocksmalltext, above = of urals] (exrate) {USD/RUB};
    \node[blocksmalltext, above = of fedbudg] (i) {RUONIA};
    \draw[<->] (urals.north) -- (exrate.south);
    \draw[<->] (urals.east) -- (fedbudg.west);
    \draw[<->] (i.south) -- (fedbudg.north);
    \draw[<->] (exrate.east) -- (i.west);
    \draw[<->] (exrate.south east) -- (fedbudg.north west);
    \draw[<->] (i.south west) -- (urals.north east);
    \node[textsmall, above = -0.05 cm of $(exrate.north)!0.5!(i.north)$, align=center] (global) {Global variables};
    \draw[thick] ($(urals.north west) +(-0.2,2.6)$) rectangle ($(fedbudg.south east) + (0.2,-0.2)$);
    \draw[->] ($(urals.west) +(-0.2,0)$) -| node[near start, above, xshift=-0.05cm] {Common shocks} ($(tradeflows.north|-urals.west) + (0,-1)$);
    \draw[<-] ($(urals.south)!0.5!(fedbudg.south) +(0,-0.2)$) |- node[midway, below, align = center] {Government reaction function} ($(rN.east) + (0.2,0)$);   
    \node[above = -0.25 cm of global] {\circled{3}};
\end{tikzpicture}
```

## Data

- **Monthly** data for 79 Russian regions:

    - Reflects short outside lag of fiscal transmission
    - Allows to justify the identification strategy

- Three variable blocks:

    1. **Regional variables**: logs of output index and expenditures of the consolidated regional budget in per capita terms, log-difference of the regional CPI
    2. **External variables**: weighted combination of regional variables for neighbouring regions. Weight matrix is constructed based on interregional trade for 2016
    3. **Global variables**: RUONIA rate, logs of Urals price, USD/RUB exchange rate, per capita revenue and expenditure of the federal budget

## GVAR Specification

Endogenous variables in region $i$ evolve according to a VAR:

$$
x_{it} = a_{i0} + a_{i1} t + \sum_{l=1}^{p_i} \Phi_{il} x_{i,t-l} + \sum_{l=0}^{q_i} \Lambda_{il} x^*_{i,t-l} + \sum_{l=0}^{s_i} D_{il} \omega_{t-l} + \varepsilon_{it}
$$

- $x_{it}$ - vector of endogenous variables
- $x^*_{i,t}$ - vector of external variables
- $\omega_t$ - vector of global variables
- $p_i$, $q_i$ and $s_i$ determine the lag order (1 in the baseline specification)

Global variables are modelled in a separaet dominant-unit VAR and react to regional economic dynamics:

- Interest rate is determined according to a Taylor rule variant which includes weighted regional output and inflation
- Federal budget revenues depend on weighted regional output, oil prices and the exchange rate

## Estimation

- Each regional model has 39 coefficients and 9 covariance matrix elements $\rightarrow$ "curse of dimensionality"

- We opt for a Bayesian estimation using the Minnesota prior. The prior assumes that the explanatory power is higher:

    - For own lags compared to other variable lags
    - For more recent lags than for more distant lags
    
- Important assumption: each region is a small open economy [@pesaranModelingRegionalInterdependencies2004]:

    - External variables are treated as weakly exogenous
    - Regional models are estimated separately
    - Global model solution is obtained algebraically using the weighting matrix
    
- The model is estimated with stochastic volatility to capture periods of markedly high volatility (2014-2015 and 2022)

## Identification Strategy

- Reduced-form VAR residuals cannot be interpreted in a structural manner since they are correlated with each other

- One needs to impose restrictions on the covariance matrix and perform orthogonalization

- Baseline identification strategy is recursive:
  
    - Regional supply and demand shocks do not influence government expenditure within one month
    - Regional budget expenditure can have a contemporaneous impact on output and inflation

- Main results are qualitatively robust to using sign restrictions as an alternative identification strategy

- N.B. Identification in GVAR models is local: the shock is orthogonal to other regional shocks, but might be correlated with shocks in other regions

    - In practice this correlation is small
    
## Government Spending Multiplier

- Regions are very different in their size and budget spending $\rightarrow$ IRFs have to be transformed for comparability

- The government spending multiplier is defined as a percentage change in output after a government spending shock equivalent to 1% of the regional per capita GRP:

$$
mult^y_{ijh} = IRF^{y,g}_{ijh} \times \frac{1}{\bar{g}_j} \times \overline{GRP}_j \times 0.01
$$

- $IRF^{y,g}_{ijh}$ - IRF of output in region $i$ to a government spending shock in region $j$ at horizon $h$

- $\bar{g}_j$ - average per capita government expenditure in region $j$

- $\overline{GRP}_j$ - average per capita GRP

## Regional Government Spending Multiplier

```{r OwnYDynReg, fig.width=9, dpi=170}
multY %>%
  ggplot() +
  geom_line(aes(x=h, y=value, group = OKATO_id), linewidth = 0.65, linetype = 'dotted', colour = 'gray') +
  geom_line(data = tableY %>%
              filter(type == 'Weighted Mean') %>%
              pivot_longer(-type) %>%
              mutate(name = as.factor(name)),
            aes(x=name, y=value, group = 1, linetype = 'dashed'), colour = 'black', linewidth = 1.25) +
  geom_line(data = tableY %>%
              filter(type == '50%') %>%
              pivot_longer(-type) %>%
              mutate(name = as.factor(name)),
            aes(x=name, y=value, group = 2, linetype = 'solid'), colour = 'black', linewidth = 1.25) +
  scale_linetype_identity(breaks = c('solid', 'dashed'),
                          labels = c("Median", "Weighted Average"),
                          guide = "legend") +
  guides(linetype=guide_legend(ncol=2)) +
  theme_hse() +
  labs(x = 'Horizon',
       y = '%',
       linetype = "Guide name"
       # title = "Экономические показатели российских регионов",
       # caption = "Источник: Росстат, Федеральное Казначейство, расчёты автора"
  ) + 
  ylim(c(tableY[1,-1] %>% min(), tableY[5,-1] %>% max())) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 18),
    axis.text = element_text(size = 20),
    axis.title.y.left = element_markdown(size = 18, margin = margin(r = 0, l = 0, unit = "cm")),
    axis.title.x = element_markdown(size = 18, angle = 0, vjust = 1.05, margin = margin(r = 0, l = 0, unit = "cm")),
    legend.key.size = unit(3,"line")
    # plot.title = element_markdown(lineheight = 1.1, size = 20),
    # plot.subtitle = element_markdown(lineheight = 1.1, size = 18),
    # plot.caption = element_markdown(family = "HSE Sans", lineheight = 1.1, hjust = 1, size = 16, face = "italic"),
    # # aspect.ratio = 0.56,
    # strip.text = element_markdown(size = 16,lineheight = 0.4),
    # strip.background = element_rect(fill = 'white'),
    # panel.border = element_rect(colour = "black", fill = NA)
  )
```


## Sources of Heterogeneity: Potential Channels

- Liquidity constraints for households and firms [@bernardiniHeterogeneousGovernmentSpending2020;@debuque-gonzalesLocalFiscalMultipliers2021]

- The source of financing and regional debt dynamics [@chodorow-reichGeographicCrossSectionalFiscal2019]

- Govrenment spending composition and allocation between labour- and capital-intensive industries [@juarrosHeterogeneousSpendingHeterogeneous2023]

- Trade linkages: spillover effects are stronger in regions with a trade deficit [@bernardiniHeterogeneousGovernmentSpending2020; @duporRegionalConsumptionResponses2023]


## Sources of Heterogeneity: Hypotheses

1. Size and structure of the economy: real GRP per capita, share of mining, manufacturing and construction in GRP
1. Income distribution: decile coefficient and share of population with incomes below the poverty line
1. Regional labor market: unemployment rate, number of unemployed per vacancy, share of employment in non-state-owned organizations, share of employment in small and medium-sized enterprises
1. Trade relations: share of imports from abroad in GRP, ratio of shipment of products from other regions of Russia and to other regions of Russia to total regional output
1. Regional budget: level of regional debt to GRP (excluding budget loans), share of regional budget expenditures on national defense, security, national economy and social sphere

## Main Conclusions: Regional Multipliers

- Median multiplier reaches `r YMax`% after `r YMaxHor` and dies out gradually

- The effect is stronger in regions with:

    - High GRP per capita (market size and lower propensity to import)
    - High share of construction in GRP
    - High income inequality (population groups with a high propensity to consume)
    - Predominance of employment in large enterprises (structure of budget recipients)
    - High share of expenditure on the national economy in the budget (productive expenditure)

## Spillover Effects: Definition 

We calculate two measures of spillover effects:

1. **Spill-out effect** reflects how strongly other regions react to a government spending shock in a given region:

$$
SO_j = \sum_{i\neq j} \omega_{i} mult^y_{ijh}
$$

2. **Spill-in effect** reflects the sensitivity of a given region to government spending shocks in other regions:

$$
SI_i = \sum_{j\neq i} \omega_j mult^y_{ijh}
$$


The responses are weighted using regional GRP: $\omega_i=\frac{GRP_i}{\sum_k GRP_k}$

## Main Conclusions: Spillover Effects

- The average spill-out effect at a 1-year horizon is `r tableYOut[6,6] %>% as.numeric() %>% round(.,2)`%. A stronger response is found in:

    - Larger economies
    - Regions with a lower share of intra-Russia exports
    - Regions with a higher share of foreign imports
    - Regions with a higher unemployed-vacancy ratio
    
- The average spill-in effect at a 1-year horizon is `r tableYIn[6,6] %>% as.numeric() %>% round(.,2)`%. A stronger response is found in:

    - Regions with a higher share of manufacturing in GRP
    - Regions with lower income inequality
    - Regions with a lower unemployed-vacancy ratio

## Discussion and Conclusions

- The results are within the range of existing estimates, but higher than the national multiplier for Russia:

    - Regional budgets vs. federal budget: expenditure structure
    - General equilibrium effects at the national level: Central Bank response, inflation expectations, debt trajectory
    - Dependent variable: GDP vs. monthly GRP proxy (covers from 30% to 80% of GVA in regions)
    
- The multiplier is higher in larger regions which can systematically conduct looser fiscal policy due to a higher and more stable tax base:

    - A join fiscal stimulus in top-10 richest regions slightly increases intraregional Gini coefficient (the effect is mediated by spillovers)
    
- Price level reaction to a government spending shock: significantly negative in approximately 50% of the cases

    - Contradicts traditional views but is often encountered empirically [@jorgensenInflationResponseGovernment2022]
    - The importance of supply-side fiscal policy effects [@jorgensenInflationResponseGovernment2022; @choiAreGovernmentSpending2022]
    
## Sources {.allowframebreaks}

\fontsize{6pt}{7.2}\selectfont

::: {#refs .blah}
:::

<!-- ## Пространственные взаимосвязи между переменными {.fragile} -->

<!-- ```{r Crosscorr, tab.id = "Crosscorr", tab.cap='Средние коэффициенты корреляции между переменными и остатками модели в регионах России'} -->
<!-- summ$crosscorrdata[,1:3] %>% -->
<!--    as.data.frame(.) %>% -->
<!--    rownames_to_column() %>% -->
<!--    left_join(summ$crosscorr[,1:3] %>% -->
<!--    as.data.frame(.) %>% -->
<!--    rownames_to_column(), by = 'rowname') %>% -->
<!--    relocate(exp.y, .before = y.x) %>% -->
<!--    relocate(y.y, .before = pi.x) %>% -->
<!--    kbl(col.names = c("Коэффициент корреляции", -->
<!--                            "Данные", -->
<!--                            "Остатки модели", -->
<!--                            "Данные", -->
<!--                            "Остатки модели", -->
<!--                            "Данные", -->
<!--                            "Остатки модели"),  -->
<!--        digits = 2, booktabs = T) %>% -->
<!--   add_header_above(c(" " = 1, "Госрасходы" = 2, "Выпуск" = 2, "Инфляция" = 2)) %>% -->
<!--   kableExtra::footnote(general = "в таблице приведено число и доля регионов, для которых средний коэффициент корреляции между региональными переменными (или остатками уравнений модели) лежит в указанном диапазоне", general_title = "Примечание: ", footnote_as_chunk = T) -->
<!-- ``` -->


## Regional Government Spending Multiplier {.fragile}

```{r OwnY, tab.id = "OwnY", tab.cap = 'Regional Government Spending Multiplier in Russia'}
tableY %>%
  filter(!type %in% c('Weighted Median'
                      ,'CU_l','CU_u'
                      )
         )%>%
  mutate(type = c('Min',
                  '15th percentile',
                  'Median',
                  '85th percentile',
                  'Maximum',
                  paste0('Average', footnote_marker_number(1, "latex")),
                  paste0('Share of significant', footnote_marker_number(2, "latex"))
                  #,'Средняя нижняя граница 68% интервала',
                  #'Средняя верхняя граница 68% интервала'
                  )) %>%
  select(`Статистика`=type, `0`, `1`, `3`, `6`, `12`, `24`) %>%
  kbl(digits = 2, escape = F, format = "latex", booktabs = T) %>%
  kableExtra::kable_styling(font_size = 7) %>%
  add_header_above(c(" " = 1, "Горизонт (месяцев)" = 6))  %>%
  kableExtra::footnote(general = 'the estimates are using a GVAR model on the sample of 2009-2024 using a recursive identification scheme. The multiplier is understood as the percentage change in the business activity index in response to an increase in government spending equivalent to 1% of GRP.', general_title = 'Note: ', number = c('When calculating the average, regional per capita GRP is used as weights', '68% credible intervals are used to test significance.'), threeparttable = T)
```

## Heterogeneity Regressions I {.fragile}

```{r Yregress, tab.id = "Yregress", tab.cap = 'Regional Fiscal Multiplier Heterogeneity'}
hor <- 3
modelData <- horizonY %>%
  filter(h == hor) %>%
  rename(y = value) %>%
  select(-c(sig68,grp)) %>%
  left_join(explanatory %>% select(-c(OKATO,sport,state,culture,utilities)))
# GRP and Structure
m1 <- lm(y ~ log(GRP) + extraction + industry + construction, modelData,
             weight = GRP)
m1r <- coeftest(m1, vcov = vcovHAC, save = TRUE)
# Liquidity constraints
m2 <- lm(y ~ decile_coeff + poverty, modelData,
             weight = GRP)
m2r <- coeftest(m2, vcov = vcovHAC, save = TRUE)
m2r <- coeftest(m2, vcov = vcovHAC, save = TRUE)
# Labour market
m3 <- lm(y ~ tightness + unemprate + private_empl_share + sme_empl_share, modelData,
             weight = GRP)
m3r <- coeftest(m3, vcov = vcovHAC, save = TRUE)
# Trade
m4 <- lm(y ~ imgrp + exsh + imsh, modelData,
             weight = GRP)
m4r <- coeftest(m4, vcov = vcovHAC, save = TRUE)
# Budget
m5 <- lm(y ~ defense + security + economy + social + debtgrp, modelData,
             weight = GRP)
m5r <- coeftest(m5, vcov = vcovHAC, save = TRUE)
# Table
tab <- huxreg("GRP" = m1, "Inequality" = m2, "Labour market" = m3, number_format = 2, statistics = c("R^2 adj." = "adj.r.squared",
      "P value" = "p.value"), bold_signif = 0.1, stars = c(`*` = 0.1, `**` = 0.05, `***` = 0.01))
tab <- tab[-seq(3,23,2),]
tab %>%
  set_font_size(5) %>%
  set_bottom_padding(0) %>%
  set_top_padding(0) %>%
  set_width(1) %>%
  set_col_width(col = 1, 0.3) %>%
  set_all_borders(brdr(0.5, "solid")) %>%
  set_align(col = 2, value = 'center') %>%
  set_align(col = 3, value = 'center') %>%
  set_align(col = 4, value = 'center')
```

## Heterogeneity Regressions II {.fragile}

```{r Yregress2, tab.id = "Yregress2", tab.cap = 'Regional Fiscal Multiplier Heterogeneity'}
tab <- huxreg("External trade" = m4, "Budget" = m5, number_format = 2, statistics = c("N" = "nobs",
      "R^2 adj." = "adj.r.squared",
      "P value" = "p.value"), bold_signif = 0.1, stars = c(`*` = 0.1, `**` = 0.05, `***` = 0.01))
tab <- tab[-seq(3,19,2),]
tab %>%
  set_font_size(5) %>%
  set_bottom_padding(0) %>%
  set_top_padding(0) %>%
  set_width(1) %>%
  set_col_width(col = 1, 0.3) %>%
  set_all_borders(brdr(0.5, "solid")) %>%
  set_align(col = 2, value = 'center') %>%
  set_align(col = 3, value = 'center')
```

## Spillover Effects: Summary Statistics

```{r SpillY, tab.id = "SpillY", tab.cap = 'Government spending spillover Effects in Russian regions'}
tableYOut %>%
  select(c(1,2,5,6,7)) %>%
  left_join(tableYIn %>% select(c(1,2,5,6,7)), by = 'Статистика', suffix = c('','.')) %>%
  mutate(`Статистика` = c('Min',
                  '15th percentile',
                  'Median',
                  '85th percentile',
                  'Maximum',
                  paste0('Average', footnote_marker_number(1, "latex")),
                  paste0('Share of significant', footnote_marker_number(2, "latex"))
                  )) %>%
  kbl(digits = 2, escape = F, booktabs = T,
      col.names = c('Statistic', 
                    '0', '6', '12', '24',
                    '0', '6', '12', '24')) %>%
  kableExtra::kable_styling(font_size = 7) %>%
  add_header_above(c(" " = 1, "Spill-out" = 4, "Spill-in" = 4)) %>%
  add_header_above(c(" " = 1, "Horizon (months)" = 8)) %>%
  kableExtra::footnote(general = "the estimates are obtained using a GVAR model on the 2009-2024 sample using a recursive identification scheme. In all cases, the shock size is normalized to 1% of the GRP of the shock source region.", general_title = "Note: ", footnote_as_chunk = T, threeparttable = T) 
```

## Interregional Inequality

```{r IneqDyn, fig.cap='Estimated Gini coefficient for per capita GRP', dpi=170, fig.height=5}
rr <- readRDS('../Data/explanatory.RDS') %>%
  select(OKATO, OKATO_id, GRP) %>%
  arrange(GRP) %>%
  left_join(codes %>% select(region = Abbrev, name_rus = `Rus Name`, name_eng = Name, OKATO_id)) %>%
  filter(!is.na(region)) %>%
  arrange(GRP)
poor10 <- rr$region[1:10]
rich10 <- rr$region[70:79]

regineq <- read_excel('../Data/explanatory.xlsx', sheet = 'Ineq') %>%
  filter(OKATO_id %in% grp$OKATO_id) %>%
  select(-region) %>%
  left_join(codes %>% select(OKATO_id, region = Abbrev)) %>% 
  relocate(region, .before = OKATO_id)
g <- apply(regineq[,-c(1:4)] %>% as.matrix(), 2, DescTools::Gini, weights = regineq$population, na.rm=T)
# richest
# with spillovers
grps <- regineq %>%
  select(region, population, `2022`) %>%
  arrange(region)
mults <- readRDS(paste0(path,'/',MODEL,'-joint_10richest_chol.RDS')) %>%
    filter(variable == 'y') %>%
    ungroup() %>%
    select(region, h, Q50) %>%
    pivot_wider(names_from = 'h', values_from = 'Q50') %>%
    arrange(region)
sim <- as.vector(grps[,3]) * (1 + mults[,-c(1)]/100)
sim <- cbind(region = grps$region, population = grps$population, sim)
giniR <- apply(sim[,-c(1:2)] %>% as.matrix(), 2, DescTools::Gini, 
                weights = sim$population, 
                na.rm=T)
# without spillovers
mults2 <- mults
mults2[which(!mults$region %in% rich10),-1] <- 0 
sim2 <- as.vector(grps[,3]) * (1 + mults2[,-c(1)]/100)
sim2 <- cbind(region = grps$region, population = grps$population, sim2)
giniR2 <- apply(sim2[,-c(1:2)] %>% as.matrix(), 2, DescTools::Gini, 
                weights = sim2$population, 
                na.rm=T)
# poorest
mults <- readRDS(paste0(path,'/',MODEL,'-joint_10poorest_chol.RDS')) %>%
    filter(variable == 'y') %>%
    ungroup() %>%
    select(region, h, Q50) %>%
    pivot_wider(names_from = 'h', values_from = 'Q50') %>%
    arrange(region)
sim <- as.vector(grps[,3]) * (1 + mults[,-c(1)]/100)
sim <- cbind(region = grps$region, population = grps$population, sim)
giniP <- apply(sim[,-c(1:2)] %>% as.matrix(), 2, DescTools::Gini, 
                weights = sim$population, 
                na.rm=T)
# without spillovers
mults2 <- mults
mults2[which(!mults$region %in% poor10),-1] <- 0 
sim2 <- as.vector(grps[,3]) * (1 + mults2[,-c(1)]/100)
sim2 <- cbind(region = grps$region, population = grps$population, sim2)
giniP2 <- apply(sim2[,-c(1:2)] %>% as.matrix(), 2, DescTools::Gini, 
                weights = sim2$population, 
                na.rm=T)
# plot
ineqDf <- data.frame(h=0:24, gini = c(giniP, giniR, giniR2), group = c(rep('Scenario 2', 25), rep('Scenario 1', 25), rep('Scenario 1 (no spillovers)', 25)))
ineqDf %>%
  ggplot() +
  geom_line(aes(x=h, y=gini, group=group, linetype = group), linewidth = 1.25) +
  geom_hline(yintercept = g['2022'] %>% round(.,3), colour = 'black', linewidth = 1) +
  #annotate("text", x=15,y=g['2022'],label = paste0('Уровень 2022 года (', g['2022'] %>% round(.,3), ')'), vjust = -1, family = "Times New Roman") +
  scale_linetype_manual(values = c("Scenario 1" = "dashed", "Scenario 1 (no spillovers)" = "dotdash", "Scenario 2" = "dotted"),
                          guide = "legend") +
  theme_hse() +
  guides(linetype=guide_legend(ncol=3)) + 
  labs(x = 'Horizon, months',
       y = NULL,
       linetype = "Guide name"
       # title = "Экономические показатели российских регионов",
       # caption = "Источник: Росстат, Федеральное Казначейство, расчёты автора"
  ) + 
  #ylim(c(g['2022']-0.002, g['2022']+0.01)) +
  scale_y_continuous(breaks = seq(from = 0.312, to = 0.32, by = 0.001)) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 18),
    axis.text = element_text(size = 20),
    axis.title.y.left = element_markdown(size = 18, margin = margin(r = 0, l = 0, unit = "cm")),
    axis.title.x = element_markdown(size = 18, angle = 0, vjust = 1.05, margin = margin(r = 0, l = 0, unit = "cm")),
    legend.key.size = unit(1.5,"line")
    # plot.title = element_markdown(lineheight = 1.1, size = 20),
    # plot.subtitle = element_markdown(lineheight = 1.1, size = 18),
    # plot.caption = element_markdown(family = "HSE Sans", lineheight = 1.1, hjust = 1, size = 16, face = "italic"),
    # # aspect.ratio = 0.56,
    # strip.text = element_markdown(size = 16,lineheight = 0.4),
    # strip.background = element_rect(fill = 'white'),
    # panel.border = element_rect(colour = "black", fill = NA)
  )
```

Scenario 1: fiscal expansion in 10 regions with the highest per capita GRP, scenario 2 - in 10 regions with the lowest per capita GRP. The bold line indicates the value of the Gini coefficient in 2022 (0.315)